---
title: "Monte Carlo Simulations"
author: "Nick Harbeck"
date: "April 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##First Steps

The first thing that we'll look at is a fake stock market trend[^1]:

```{r stocks)}
days <- 200
changes <- rnorm(200,mean=1.002,sd=0.005)
plot(cumprod(c(19,changes)),type='l',ylab="Price",xlab="day",main="Make that Money $$$")
```

Then Monte Carlo simulations can be used to predict what will happen with this stock over the course of many runs
```{r trials}
runs <- 100000
#simulates future movements and returns the closing price on day 200
generate.path <- function(){
  days <- 200
  changes <- rnorm(200,mean=1.002,sd=0.005)
  sample.path <- cumprod(c(19,changes))
  closing.price <- sample.path[days+1] #+1 because we add the opening price
  return(closing.price)
}

mc.closing <- replicate(runs,generate.path())
```

You can then estimate the probabilities for the ending value of the stock by analyzing mc.closing

quantile(mc.closing,0.95) = 27.35

quantile(mc.closing,0.05) = 21.69

[^1]: Sourced from https://www.countbayesie.com/blog/2015/3/3/6-amazing-trick-with-monte-carlo-simulations

##Other Examples

###Setting up a simulation
The MonteCarlo package in r is a good tool for exploring these simulations and I recommend exploring the documentation [^2]

Exploring further, let's look at some historical data from a data set to start our Monte Carlo simulation. 
```{r example}
mean(mtcars$mpg)
sd(mtcars$mpg)

simulatedMPG <- rnorm(n = 1000, mean = mean(mtcars$mpg), sd = sd(mtcars$mpg)) #number of simulated values can be set to any number

#You can also change the distribution type in the simulation which is powerful for distributions that are not normal

#Now to take a look at the sample
head(simulatedMPG) #These are the first five randomly sampled results from the simulatedMPG distribution.
mean(simulatedMPG) 
sd(simulatedMPG)
#As expected, the mean and standard deviation of the simualtedMPG distribution are close to the historical dataset.

#This simulation can be replicated

simulatedMPGrep <- replicate(n=5, simulatedMPG)
head(simulatedMPGrep) #First five results for each replication
summary(simulatedMPGrep)
```

You can then perform statistical analysis of the simulations from above to see how the experiments approach the mean and standard deviation that you expect

```{r wildmetaanalysis}
mean(simulatedMPGrep)
sd(simulatedMPGrep)
```

###There are more options for Monte Carlo Simulations besides distribution simulations
There's a good series of tutorials on YouTube for regressions [^3]

[^3]: https://www.youtube.com/watch?v=xuUMz8exU8Q

Now for some optimization and finding a minimum value:

```{r optimization}
RandomFunction <- function(x){ 1/7*(((1000+x^2)^1/2))+1/3*(((1000+(50-x))^1/2))}

RandomFunction(3) #Feel free to plug in a number here to find an output

plot(RandomFunction(-200:400))
```

Now that a function is defined, Monte Carlo simulations can estimate the minimum.

```{r MonteOpt}
EstimateX <- runif(1000, min = -200, max = 400) #This is a uniform distibution of 1000 random X values from -200 to 400

head(RandomFunction(EstimateX))
summary(RandomFunction(-200:400)) #actual summary values
summary(RandomFunction(EstimateX)) #Estimates using a Monte Carlo simulation
```

This can be applied to other functions and is also useful for finding local minimums or maximums by adjusting bounds in the uniform distribution.





[^2]: https://cran.r-project.org/web/packages/MonteCarlo/vignettes/MonteCarlo-Vignette.html